---
title: "Dissertation - WhatsMatching"
author: "Jason Meyer, z5242267"
date: "2022-11-18"
output: 
  pdf_document:
    latex_engine: xelatex
bibliography: vignettes/App_References.bib
csl: vignettes/ama.csl
link-citations: yes
linkcolor: blue
nocite: '@*'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Preface {.unnumbered}

This project has been formally titled "What's Matching". This document represents the formal submission via *turnitin* for the dissertation component of the Master of Health Data Science Degree at UNSW (HDAT9900-HDAT9901-HDAT9902-HDAT: Dissertation). The full and properly formatted outputs for the assignment can be accessed in the following places:\
- [WhatsMatching Website](https://jmeyer2482.github.io/WhatsMatching)\
- [WhatsMatching Shiny App](https://jmeyer2482.shinyapps.io/WhatsMatching)\
- [WhatsMatching Github Repository](https://github.com/jmeyer2482/WhatsMatching)

```{r child='index.md'}

```

```{r child='vignettes/Background.rmd'}

```

```{r child='vignettes/Development.rmd'}

```

```{r child='vignettes/TheApp.rmd'}

```

```{r child='vignettes/DataGen.rmd'}

```

```{r child='vignettes/Matching.rmd'}

```

```{r child='vignettes/Examples.rmd'}

```

```{r child='vignettes/Summary.rmd'}

```

## References
