---
title: "PSM ShinyApp Plan"
author: "Jason Meyer"
date: "25/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(MatchIt)
library(cem)
library(ggpubr)
library(optmatch)
library(cobalt)
library(Hmisc)
library(tidyverse)
```

# Shiny app for the exploration of the PSM paradox

Take simulated or pre-existing `R` data sets and use them to generate graphical outputs that provide intuitive insights into how the propensity operates.

Available data sets for this project will use `lalonde` and `nuclearplants` which are used as example data for matching in the `MatchIt` and `optmatch` packages respectively. Additionally, the `FEV` data set from [Vanderbilt Biostatistics Datasets](https://hbiostat.org/data/) has also been included via the `Hmisc` package. 


```{r, FEV, include=FALSE}

## quick review of FEV
# getHdata(FEV)
# 
# 
# FEV.match <- matchit(smoke~age+height+sex, FEV, distance = "mahalanobis")
# 
# match.out <- match.data(FEV.match)
# 
# 
# lm(fev~smoke, match.out) %>% summary()
# lm(fev~smoke+age+height+sex, FEV) %>% summary()

```


## Simulated data

```{r}

#normal distribution
rnorm(200, 27, 5.5) %>% hist(main="Normal Distribution")

#uniform distribution
runif(200, 10, 30) %>% hist(main="Uniform Distribution")

#poisson distribution
rpois(200, 25) %>% hist(main="Poisson Distribution")

#binomial distribution
rbinom(200, 20, 0.5) %>% hist(main="Binomial Distribution")

#beta 
rbeta(200, 10, 1) %>% hist(main="Beta Distribution")

rcauchy(200)
```


```{r}

golem::create_golem("../PSM.golex")

```

