---
title: "Vignette 2 - The App"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette 2 - The App}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# library(WhatsMatching)
```


## The nuts and bolts (“the server side”)  

This project has been developed as a Shiny app using the Golem framework to support it's production. There are a number of key functions that have been implemented to generate the data and matching processes. Let's have a look in a bit more detail. 

### The golem framework  

The <a href="https://golemverse.org/" target="_blank">Golem framework</a> is the brain child of Colin Fay. In essence, Shiny apps are ideal for development as a R package. That is to say that, the package structure lends itself well to the development of production grade Shiny apps.  

The `golem` package itself provides functions and helpers to generate and create a framework that can be used for the development of the Shiny app. In addition, there are theoretical underpinnings and best practice advice provided to give the best chance of developing a high quality application.

### Key functions and functionality  

There are <a href="https://jmeyer2482.github.io/WhatsMatching/reference/index.html" target="_blank">four key functions</a> that have been developed as part of the package in order to make data generation and analysis as straightforward as possible. They are `create.sim.data`, `matched.data`, `matching.plot`, and `combined.plot`. The functions have been designed to follow and iterative process.

 - `create.sim.data()`


[WhatsMatching::create.sim.data()]  


### Deployment (The app is here, the code is here and the vignettes are here)  

## Simulations (“the UI side”)  

### Example A (data | user options | screenshot)  

### Example B (data | user options | screenshot)  

### Example C (data | user options | screenshot)  

### Example D (data | user options | screenshot)  

### Example E (data | user options | screenshot)  


## Introduction  
  
  This Shiny App package has been developed with the intention of providing educational support for understanding matching methods in the context of causal inference. The initial intention was to deal specifically with the Propensity Score Paradox highlighted by King and Neilsen[ref]. Over the development of the app we decided that the underlying theoretical concepts that inform how matching is conducted by might be more useful and engaging in an educational environment. This document outlines the theoretical concepts and development process that was undertaken to develop the app.  
  
## Matching Visuals  
  
  In order to demonstrate how matching works visually, we will simulate some data under different conditions to highlight the differences in the methods. This will purely show how data are matched between groups, we will look out how that changes the estimate of the treatment effect later.
  
  All these simulations will contain 60 control units and 40 treated units and the matches will be completed with both the Propensity Score and the Mahalanobis Distance. 
  
  
```{r simulated data}



# 
# #treatment has a mild effect on one covariate (1/2 std dev)
# d$X1.mild <- d$t*2.5 + d$X1
# 
# #treatment has a strong effect on one covariate (3 std dev)
# d$X1.strong <- d$t*15 + d$X1
# 
# #treatment has opposite effects on covariates (mild)
# d$X2.mild <- d$t*-2 + d$X2
# 
# #treatment has opposite effects on covariates (strong)
# d$X2.strong <- d$t*-12 + d$X2




```
  
  
```{r}
# #random, no treatment effect
# psm.1 <- WhatsMatching::matched.data(t~X1+X2, d, "Propensity Score")
# mdm.1 <- WhatsMatching::matched.data(t~X1+X2, d, "Mahalanobis")
# 
# psp.1 <- plot.matches(psm.1, "Propensity Score Matches - Random Data")
# mdp.1 <- plot.matches(mdm.1, "Mahalanobis Distance Matches - Random Data")
# 
# 
# psm.2 <- WhatsMatching::matched.data(t~X1.mild+X2, d, "Propensity Score")
# mdm.2 <- WhatsMatching::matched.data(t~X1.mild+X2, d, "Mahalanobis")
# 
# psp.2 <- plot.matches(psm.2, "Propensity Score Matches - Mild treatment effect on X1")
# mdp.2 <- plot.matches(mdm.2, "Mahalanobis Distance Matches - Mild treatment effect on X1")
# 
# 
# psm.3 <- WhatsMatching::matched.data(t~X1.strong+X2, d, "Propensity Score")
# mdm.3 <- WhatsMatching::matched.data(t~X1.strong+X2, d, "Mahalanobis")
# 
# psp.3 <- plot.matches(psm.3, "Propensity Score Matches - Strong treatment effect on X1")
# mdp.3 <- plot.matches(mdm.3, "Mahalanobis Distance Matches - Strong treatment effect on X1")
# 
# 
# psm.4 <- WhatsMatching::matched.data(t~X1.mild+X2.mild, d, "Propensity Score")
# mdm.4 <- WhatsMatching::matched.data(t~X1.mild+X2.mild, d, "Mahalanobis")
# 
# psp.4 <- plot.matches(psm.4, "Propensity Score Matches - Mild treatment effect on X1 and X2")
# mdp.4 <- plot.matches(mdm.4, "Mahalanobis Distance Matches - Mild treatment effect on X1 and X2")
# 
# 
# psm.5 <- WhatsMatching::matched.data(t~X1.strong+X2.strong, d, "Propensity Score")
# mdm.5 <- WhatsMatching::matched.data(t~X1.strong+X2.strong, d, "Mahalanobis")
# 
# psp.5 <- plot.matches(psm.5, "Propensity Score Matches - Strong treatment effect on X1 and X2")
# mdp.5 <- plot.matches(mdm.5, "Mahalanobis Distance Matches - Strong treatment effect on X1 and X2")
# 
# 
# 
# # ggpubr::ggarrange(psp.1, mdp.1, common.legend = TRUE)
# # ggpubr::ggarrange(psp.2, mdp.2, common.legend = TRUE)
# # ggpubr::ggarrange(psp.3, mdp.3, common.legend = TRUE)
# # ggpubr::ggarrange(psp.4, mdp.4, common.legend = TRUE)
# # ggpubr::ggarrange(psp.5, mdp.5, common.legend = TRUE)
# 
# ggpubr::ggarrange(psp.1,psp.2,psp.3, mdp.1, mdp.2, mdp.3, common.legend = TRUE, legend = "right", align = "hv")
# 
# 
# ggpubr::ggarrange(psp.1,psp.4,psp.5, mdp.1, mdp.4, mdp.5, common.legend = TRUE, legend = "right")

```
  
  
  
  
  
  
  
  For the purposes of simplicity and uniformity the `optmatch` package has been used to generate a distance matrix using the `match_on` function. The format of the data that is returned is the same for both the Mahalanobis Distance and the Propensity Score. We'll also be use the `fev` dataset from the `mplot` package which contains a data that is useful for demonstrating the matching process.     
  
  
```{r setup2, echo=FALSE}
#function to load data directly to named variable instead of
#the variable being loaded into the global environment
# load.data <- function(dt, pkg){
#   
#   # make sure the relevant library is available
#   library(pkg, character.only = T)
#   #create environment for loading, otherwise loads into global
#   e <- new.env()
#   #get data
#   data(list=dt, package=pkg, envir = e)
#   #store data
#   d <- e[[dt]]
#   #delete environment
#   e <- NULL
#   #return data
#   return(d)
# }

fig.count <- 1
#this is a custom function for loading data into named variable
# dt <- load.data("fev", "mplot")

```
  
  Let's have a quick look at the data so we know what we're dealing with.

```{r preview data}

# head(dt)

```

  As we can see, there are 5 columns. All the data here is numeric. For the purpose of this exercise, we want to see what the effect of smoke is on forced expiratory volume. Therefore, we will use the `smoke` variable as the treatment and `fev` is the outcome. The remaining covariates are `age`, `height`, and `sex`.  
  
  Let's have a quick look at the data using `ggplot2`.  
  
```{r data plots}
# library(dplyr)
# library(ggplot2)
# 
# #add meaningful labels
# dt <- dt %>% mutate(Sex=ifelse(sex==0,"Female","Male"),
#                    Smokes=ifelse(smoke==0,"No","Yes"))
# 
# p1 <- ggplot(dt, aes(x=Smokes)) 
# 
# p1.1 <- p1 + geom_boxplot(aes(y=age, fill=Smokes)) 
# p1.2 <- p1 + geom_boxplot(aes(y=height, fill=Smokes))
# p1.3 <- p1 + geom_count(aes(y=Sex, colour=Smokes))
# p1.4 <- p1 + geom_boxplot(aes(y=fev, fill=Smokes))
# 
# 
# p2 <- ggplot(dt, aes(y=fev)) 
# 
# p2.1 <- p2 + geom_point(aes(x=age, color=Smokes), alpha=0.5) 
# p2.2 <- p2 + geom_point(aes(x=height, color=Smokes), alpha=0.5)
# p2.3 <- p2 + geom_boxplot(aes(x=Sex, fill=Smokes))
# p2.4 <- p2 + geom_boxplot(aes(x=Smokes, fill=Smokes))
```


#### Fig `r fig.count`  

```{r data plots1}
# ggpubr::ggarrange(p1.1,p1.2,p1.3,p1.4, common.legend = T)
fig.count <- fig.count + 1
```
  
#### Fig `r fig.count`  

```{r data plots2}
# ggpubr::ggarrange(p2.1,p2.2,p2.3,p2.4, common.legend = T)
fig.count <- fig.count + 1
```

  
  

  
  We can see from this quick look at the data that smokers are older, taller and appear to have a larger FEV. They are also a small subset of the overall population in the dataset. If we were to map out the relationships in a directed acyclic graph, it might look something like this:
  
#### Fig `r fig.count`

```{r dag}
# fig.count <- fig.count + 1
# 
# #smoke, age, height, sex, fev
# fev.dag <- dagitty::dagitty('dag{
#                  age -> height
#                  sex -> height
#                  age -> smoke
#                  height -> fev
#                  smoke -> height
#                  smoke -> fev
#                  age -> fev
#                  age -> smoke
#                  smoke [exposure]
#                  fev [outcome]
#                  }', layout=FALSE)
# 
# 
# dagitty::coordinates(fev.dag) <- 
#   list(x=c(age=0.5,fev=1,height=0.5,sex=0,smoke=0),
#        y=c(age=0.9,fev=0.5,height=0.1,sex=0,smoke=0.5))
# 
# plot(fev.dag)


```
  
  We can see from the relationships here that `age` is a confounder as it effects both the exposure (`smoke`) and the outcome (`fev`). `height` is also important as it is a mediator. `sex` appears to be completely accounted for by `height`. Now that we have the variables for our model, let's look at some methods for matching.   

  There is more than one way of matching units in a dataset. The two methods we are looking at here are Propensity Score Matching (PSM) and Mahalanobis Distance Matching (MDM). We will also look at some of the different specifications you can make when matching.   
   
  Propensity score matching is relatively straight forward. Using logistic regression, create a vector of probabilities based on the whether or not each unit is likely to have received the treatment (or been exposed to subject of interested). We will be using `smoke ~ age + height` to determine our matches. For consistency in output between methods, I have used the `optmatch` package to generate the distance matrix.    
  
```{r create ps matix}
# f <- smoke~age+height

#fit model
# fev.mod <- glm(f, data=dt, family="binomial") 

#get matrix using the fitted values from the model (propensity scores)
#the matrix is generated using the treatment variable (y in the model output) 
# ps.dm <- optmatch::match_on(x=fev.mod$fitted.values, z=fev.mod$y)

```

  If we wanted to do the same thing for MDM then we specify it as per the optmatch documentation. In terms of the differences between the distance matrices, the propensity score distance matrix is the distance (or difference) between the calculated propensity score for each unit whereas the mahalanobis distance matrix will be the actual distance between the units based on the covariance matrix of $X$.    
  
```{r create mahalanobis matrix}
#Mahalanobis distance matrix
# md.dm <- optmatch::match_on(x=f, data=dt, method="mahalanobis")

```

  Now that we have a distance matrix for each method we can look at matching.  
  
  At this point we can start to specify how the matching occurs. Options include:  
- whether or not to use replacement  
- the order the data is matched in (this is really only important if you aren't replacing matches)  
- whether or not to use a calliper (and how tight to make it)  
  
  This exercise is not looking at the use of a calliper. It is also not exploring exact matching or other more complex match settings. We will be doing a simple match without replacement, ordered by the data for both methods.  

  Each matrix is $m \times n$ where $m$ represents the treatment group and $n$ represents the control group. To get our matches, we sequentially go through each $m$ and select the $n$ with the smallest value until we get to the last $m$. When do this with replacement there is no regard for whether $n$ has been used before, without replacement mean that the same $n$ cannot be used twice. Let's get some matches!  
  
```{r get matches}
# 
# #to make this repeatable I'm just going to wrap this is a function
# match.ids <- function(dm){ 
#     
#   #we can get the row and column names from the matrix to make looping a bit easier. 
#   #It will be the same for both methods in this case.
#   t.names <- attributes(dm)$dimnames$treatment
#   c.names <- attributes(dm)$dimnames$control
#   
#   # initialise a dataframe for storing matches
#   ids <- list(treatment=character(),
#               control=character(),
#               dist=numeric()
#   ) %>% as.data.frame()
#   
#   #loop over the distance matrix rows using the order they are already in
#   for (i in t.names) {
#     #select the row
#     
#     d <- dm[i,!c.names %in% ids$control]
#     
#     c.val <- names(which.min(d))
#     if (length(d)==1) c.val <- c.names[!c.names %in% ids$control]
#     
#     a <- list(treatment = i,
#               control = c.val,
#               dist = min(d))
#     
#     #add the matched pair to the dataframe
#     ids <- rbind(ids,a)
#   
#     row.names(ids) <- 1:nrow(ids)
#   
#   }
#   
#   return(ids)
#   
# }
# 
# #we can specify @.Data to get the matrix explictly as the optmatch 
# #object has some other parts we don't require
# ps.dm <- ps.dm@.Data
# md.dm <- md.dm@.Data
# 
# 
# #the paired ids for all the matches
# ps.ids <- match.ids(ps.dm)
# md.ids <- match.ids(md.dm)
# 
# #are the matches the same?
# table(ps.ids$control==md.ids$control)
```

  OK. Now we have our matches for each method. Let's see what they look like visually using `ggplot2`.  
  
```{r plot matches}
# ps.matches <- rbind(
#   cbind(dt[ps.ids$treatment,], subclass=ps.ids$treatment),
#   cbind(dt[ps.ids$control,], subclass=ps.ids$treatment))
# 
# md.matches <- rbind(
#   cbind(dt[md.ids$treatment,], subclass=md.ids$treatment),
#   cbind(dt[md.ids$control,], subclass=md.ids$treatment))
# 
# mp1 <- ggplot(ps.matches, aes(height, age, group=subclass)) +
#   geom_point(data=dt, aes(height, age), colour="black", alpha=0.3,
#              inherit.aes = F, size=1) + 
#   geom_point(aes(color=Smokes), alpha=0.6, size=3) + 
#   geom_line(color="black") +
#   labs(title="Propensity Score Matching matches")
# 
# mp2 <- ggplot(md.matches, aes(height, age, group=subclass)) +
#   geom_point(data=dt, aes(height, age), colour="black", alpha=0.3,
#              inherit.aes = F, size=1) + 
#   geom_point(aes(color=Smokes), alpha=0.6, size=3) + 
#   geom_line(color="black") +
#   labs(title="Mahalanobis Distance Matching matches")
```


#### Fig `r fig.count`
  
```{r plot matches1}
# mp1
fig.count <- fig.count + 1
```
  
#### Fig `r fig.count`
  
```{r plot matches2}
# mp2
fig.count <- fig.count + 1
```
    
  We can see from these two plots that the matches between the groups are slightly different between methods. For example, the smoker that is `age`=15 and `height`=60 is paired with very different non-smokers. Overall the matches look quite similar in the distribution even though there are a few differences between them. Let's see if that changes the treatment effect.  
  
  There are two ways we can specify how we estimate the treatment effect on the matched data. The purpose of the matching is to control confounding and bias. If we are happy that has been done by virtue of the matching then we need only specify `fev ~ smoke`. Alternatively, if we wanted to cover our bases, we could include the covariates in the model as well - `fev ~ age + height + smoke`. Let's compare them to a model that contains all the data.  
  
```{r view model estimates}
# f1 <- fev ~ smoke
# f2 <- fev ~ height + smoke
# f3 <- fev ~ age + smoke
# f4 <- fev ~ age + height + smoke
#   
# mod1.1 <- lm(f1, md.matches) 
# mod1.2 <- lm(f1, ps.matches) 
# mod1.3 <- lm(f1, dt) 
# 
# mod2.1 <- lm(f2, md.matches) 
# mod2.2 <- lm(f2, ps.matches) 
# mod2.3 <- lm(f2, dt)
# 
# mod3.1 <- lm(f3, md.matches) 
# mod3.2 <- lm(f3, ps.matches) 
# mod3.3 <- lm(f3, dt)
# 
# mod4.1 <- lm(f4, md.matches) 
# mod4.2 <- lm(f4, ps.matches) 
# mod4.3 <- lm(f4, dt)
# 
# mod.comp <- rbind(
#   c(model="Mahalanobis Match", formula=deparse(f1), coef(mod1.1)["smoke"], confint(mod1.1)["smoke",]),
#   c(model="Propensity Score Match", formula=deparse(f1), coef(mod1.2)["smoke"], confint(mod1.2)["smoke",]),
#   c(model="Unmatched", formula=deparse(f1), coef(mod1.3)["smoke"], confint(mod1.3)["smoke",]),
#   
#   c(model="Mahalanobis Match", formula=deparse(f2), coef(mod2.1)["smoke"], confint(mod2.1)["smoke",]),
#   c(model="Propensity Score Match", formula=deparse(f2), coef(mod2.2)["smoke"], confint(mod2.2)["smoke",]),
#   c(model="Unmatched", formula=deparse(f2), coef(mod2.3)["smoke"], confint(mod2.3)["smoke",]),
#   
#   c(model="Mahalanobis Match", formula=deparse(f3), coef(mod3.1)["smoke"], confint(mod3.1)["smoke",]),
#   c(model="Propensity Score Match", formula=deparse(f3), coef(mod3.2)["smoke"], confint(mod3.2)["smoke",]),
#   c(model="Unmatched", formula=deparse(f3), coef(mod3.3)["smoke"], confint(mod3.3)["smoke",]),
#   
#   c(model="Mahalanobis Match", formula=deparse(f4), coef(mod4.1)["smoke"], confint(mod4.1)["smoke",]),
#   c(model="Propensity Score Match", formula=deparse(f4), coef(mod4.2)["smoke"], confint(mod4.2)["smoke",]),
#   c(model="Unmatched", formula=deparse(f4), coef(mod4.3)["smoke"], confint(mod4.3)["smoke",])
# ) %>% unlist() %>% as.data.frame() %>% 
#   mutate(across(smoke:`97.5 %`, as.numeric)) 
# 
# colnames(mod.comp) <- c("model", "formula", "estimate", "CI2.5", "CI97.5")
# 
# library(plotly)
# 
# ggplotly(ggplot(mod.comp, aes(x=estimate, y=paste0(model,"\n",formula), colour=model)) +
#   geom_point() + geom_errorbarh(aes(xmin=CI2.5, xmax=CI97.5), height=0.2) +
#   labs(y="Method and Formula for Producing Estimate", x="Estimate", color="Method"))

```
  





```{r}

# dagitty::localTests(fev.dag, md.matches, type="cis.chisq")

```







  
```{r}
# library(dplyr)
# library(ggplot2)
# devtools::load_all()
# 
# #function to plot matches
# plot.matches <- function(md, title, lims=NULL){
#   
#     x.var <- all.vars(md$formula)[2]
#     y.var <- all.vars(md$formula)[3]
#   
#     p <- ggplot(md$matched.data, aes(.data[[x.var]], .data[[y.var]], group=subclass)) +
#       geom_point(data=md$data, aes(.data[[x.var]], .data[[y.var]]), colour="black", alpha=0.3,
#                  inherit.aes = F, size=3) + 
#       geom_point(aes(color=factor(t)), alpha=0.6, size=5) + 
#       geom_line(color="black") + 
#       labs(title=title, color="Treated")
# 
#     if(!is.null(lims)) p <- p + xlim(lims[1]) + ylim(lims[2])
#     
#     p
# }
# 
# #number of obs
# n <- 100
# 
# #data frame with treatment and random covariates
# d <- list(t=c(rep(0,n*.6), rep(1,n*.4)),
#           X1=rnorm(n, 50, 5), 
#           X2=rnorm(n, 40, 4)
#           ) %>% as.data.frame()
# 
# m.norep <- WhatsMatching::matched.data(t~X1+X2, d, "Mahalanobis", replace = F)
# m.rep <- WhatsMatching::matched.data(t~X1+X2, d, "Mahalanobis", replace = T)
# 
# ggpubr::ggarrange(
# plot.matches(m.norep, "Mahalanobis Matches without Replacement"),
# plot.matches(m.rep, "Mahalanobis Matches with Replacement"), common.legend = T
# )

```

  
  
  
```{r}
# 
# m.data <- WhatsMatching::matched.data(t~X1+X2, d, "Propensity Score", replace = F, order = "data")
# m.small <- WhatsMatching::matched.data(t~X1+X2, d, "Propensity Score", replace = F, order = "smallest")
# m.large <- WhatsMatching::matched.data(t~X1+X2, d, "Propensity Score", replace = F, order = "largest")
# m.rand <- WhatsMatching::matched.data(t~X1+X2, d, "Propensity Score", replace = F, order = "random")
# 
# 
# m.PO <- ggplot(m.data$matched.data, aes(X1, X2, group=subclass)) +
#   geom_point(data=m.data$data, aes(X1, X2,fill=factor(t)),alpha=0.5, shape=21,
#              inherit.aes = F, size=3) + 
#   # geom_point(aes(color=factor(t)), alpha=0.6, size=5) + 
#   geom_line(aes(linetype="data", color="data"), lwd=1, alpha=0.6) + 
#   geom_line(data=m.small$matched.data, aes(linetype="smallest",color="smallest"), lwd=1, alpha=0.6) +
#   geom_line(data=m.large$matched.data, aes(linetype="largest",color="largest"), lwd=1, alpha=0.6) +
#   geom_line(data=m.rand$matched.data, aes(linetype="random",color="random"), lwd=1, alpha=0.6) +
#   labs(title="Propensity score matching using different match ordering", fill="Treated") +
#   scale_linetype_manual(name="Ordering", values = c("data"=1, "smallest"=2, "largest"=3, "random"=4)) +
#   scale_color_manual(name="Ordering", values = c("data"="black", "smallest"="blue", "largest"="darkgreen", "random"="red")) 
#   
# m.data <- WhatsMatching::matched.data(t~X1+X2, d, "Mahalanobis", replace = F, order = "data")
# m.small <- WhatsMatching::matched.data(t~X1+X2, d, "Mahalanobis", replace = F, order = "smallest")
# m.large <- WhatsMatching::matched.data(t~X1+X2, d, "Mahalanobis", replace = F, order = "largest")
# m.rand <- WhatsMatching::matched.data(t~X1+X2, d, "Mahalanobis", replace = F, order = "random")
# 
# 
# m.MO <- ggplot(m.data$matched.data, aes(X1, X2, group=subclass)) +
#   geom_point(data=m.data$data, aes(X1, X2,fill=factor(t)),alpha=0.5, shape=21,
#              inherit.aes = F, size=3) + 
#   # geom_point(aes(color=factor(t)), alpha=0.6, size=5) + 
#   geom_line(aes(linetype="data", color="data"), lwd=1, alpha=0.6) + 
#   geom_line(data=m.small$matched.data, aes(linetype="smallest",color="smallest"), lwd=1, alpha=0.6) +
#   geom_line(data=m.large$matched.data, aes(linetype="largest",color="largest"), lwd=1, alpha=0.6) +
#   geom_line(data=m.rand$matched.data, aes(linetype="random",color="random"), lwd=1, alpha=0.6) +
#   labs(title="Propensity score matching using different match ordering", fill="Treated") +
#   scale_linetype_manual(name="Ordering", values = c("data"=1, "smallest"=2, "largest"=3, "random"=4)) +
#   scale_color_manual(name="Ordering", values = c("data"="black", "smallest"="blue", "largest"="darkgreen", "random"="red")) 
# 
# 
# 
# 
# 
# 
# ggpubr::ggarrange(
#   m.PO, m.MO, common.legend = T, legend = "right"
# )
# 

```


  
  The propensity score is a popular choice for matching but using it appropriately has been a contentious topic in recent years. Let's look at why that might be. Below, some data has been simulated with both `X1` and `X2` being mediators between `t` (the treatment variable) and `y` (the outcome variable). There is a large treatment effect between `t` and `X2` which carries on to `y`. We can see on the plot below that there is complete separation between the treated and control groups with respect to these covariates.
  
  
```{r}
# # library(WhatsMatching)
# library(ggplot2)
# 
# #sample size
# n <- 200
# #treatment effect
# te <- 4
# 
# #treatment variable with less treated units
# t <- rbinom(200,1,0.4)
# #covariates
# X1 <- rnorm(n,20,5)
# X2 <- rnorm(n,10,2)
# 
# #lets make a big treatment effect on one of the covariates
# #t*10 will double the value of X2 for treated units (on average)
# X2 <- X2 + t*2
# 
# y <- t*te + X1 + X2 + rnorm(n)
# 
# dt <- cbind(t, X1, X2, y) %>% as.data.frame()
# 
# dt %>% ggplot() + geom_point(aes(X1, X2, color=factor(t)))
```


```{r}
# mod <- glm(formula=t~X1, data=dt, family="binomial")
# 
# c(var="X1", estimate=coef(mod)["X1"], confint(mod)["X1",])
# coef(mod)[["X1"]]
# 
# dt.match <- matched.data(t~X1, dt, "Mahalanobis")
# 
# get.estimates(match.data=dt.match, f=y~t+X1+X2)
# 
# lm(y~X1+X2, data=dt.match$data)
# 
# dt.match$treatment


# a <- c("smoke", "fev", "t", "y")[c("fev", "smoke", "t", "y") %in% colnames(fev)]
# 
# a <- NULL
# 
# 
# 
# paste(NULL,a, sep="+", collapse=" ")
# 
# noquote(paste("y", "~", "t", paste(NULL,NULL, sep="+ ", collapse = " ")))
# 
# 
# as.formula(t~y+X1)

```
  
  
  
